% !TeX spellcheck = en_US
\documentclass[longdoc,parskip=full,bigchapter,linedtoc,colorback,accentcolor=tud9c,type=bsc]{tudthesis}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{array}
\usepackage{url}
\usepackage[hypcap=true]{caption}
\usepackage{hyperref}
\usepackage{float}
\usepackage{cmap}
\usepackage[space=true]{accsupp}
\usepackage{textcomp}
\usepackage{newfloat}
\usepackage{calc}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{afterpage}

\newcommand\blankpage{%
	\afterpage{
	\null
	\thispagestyle{empty}%
	\newpage}}

\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}
\def\chapterautorefname{Chapter}

\graphicspath{{graphics/}}

\DeclareFloatingEnvironment[
	fileext=cex,
	listname={List of Code Extracts},
	name={Code Extract},
	placement=h,
	within=chapter
]{code}

\newcommand\noncopynumber[1]{
	\BeginAccSupp{method=escape,ActualText={}}
	#1
	\EndAccSupp{}
}

\lstset{
	basicstyle=\ttfamily,
	columns=fullflexible,
	keepspaces=true,
	numbers=left,
	xleftmargin=2.3em,
	numberstyle=\noncopynumber,
	literate={-}{-}1
	         {*}{*}1,
	upquote=true,  
	tabsize=2,
}

\begin{document}
	
	\thesistitle{Privacy Threat Analysis Of Browser Extensions}{Analyse der Privatsphäre von Browser-Erweiterungen}
	\author{Arno Manfred Krause}
	\referee{Prof. Dr. Stefan Katzenbeisser}{Dipl. Inf. Nikolay Matyunin}
	\department{Fachbereich Informatik}
	\group{Security Engineering Group}
	\makethesistitle
	
	\section*{Abstract}
	
	Browser extensions are widely used today to enhance the way end users interact with the Internet. However, the broad range of functionality that the browser provides to its extensions can be misused to harm the user and violate his privacy. In last years, a number of malicious extensions were discovered by security experts, and researchers have already developed ways to detect them. In our work we present an in-depth analysis of possible threats to the privacy of an extension's user. We show that even an extension's basic functionality can be misused to launch targeted attacks against users and violate their privacy. As a proof-of-concept, we implemented a system which integrates a range of different attack vectors into existing extensions, and which is applicable to the majority of modern browsers, such as Chrome, Firefox, Edge, and Opera. We demonstrate that existing extensions already possess the privileges that allow to integrate these attack vectors without the need for additional privileges. Finally, we discuss possible countermeasures against malicious extensions and our implementations.
	\blankpage
	
	\affidavit{Arno Krause}
	\blankpage
	
	\tableofcontents
	\blankpage
	
	\include{./chapters/introduction}
	
	\include{./chapters/related_works}
	
	\include{./chapters/background}
	
	\include{./chapters/threat_analysis}
	
	\include{./chapters/design}
	
	\include{./chapters/extension_analysis}
	
	\include{./chapters/countermeasures}
	
	\include{./chapters/limitations_further_work}
	
	\include{./chapters/conclusion}

	\appendix
	\listof{table}{List of Tables}
	\listof{figure}{List of Figures}
	\bibliographystyle{abbrv}
	\bibliography{references}
	
\end{document}